---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# selfcomplexity

<!-- badges: start -->
[![R-CMD-check](https://github.com/edinkasia/self-complexity-package/workflows/R-CMD-check/badge.svg)](https://github.com/edinkasia/self-complexity-package/actions)
<!-- badges: end -->

Literature on self-complexity uses a variety of indices to calculate how complex, or consistent, people's selves are. This package offers functions that automate the calculation of the most frequently used indices, including Overlap, H-index, etc.

## Installation

You can install the released version of selfcomplexity from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("edinkasia/self-complexity-package")
```

## Quick demo

Use the `calculate_overlap()` function to create a dataframe with overlap scores for an example dataset:

```{r example}
library(selfcomplexity)
data(minimal_example, package = "selfcomplexity")
overlap <- calculate_overlap(data = minimal_example, att_column = "Attributes",
                             id_column = "ResponseId", subtype_column = "Subtype_name",
                             na_name_rm = TRUE)
head(overlap)
```


