---
title: "selfcomplexity"
output: rmarkdown::html_vignette
description: >
  This file describes the functions provided by selfcomplexity and the example 
  datasets available in the package.
vignette: >
  %\VignetteIndexEntry{selfcomplexity}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(selfcomplexity)
```

Researchers interested in the structure of the self-concept often use the Showers
card sort task, where they ask participants to list their self-aspects and then
identify attributes that describe each aspect of their self. The outcome of the
card sort is then used as a basis to calculate various indices that quantify 
the structure of the self. This package has been built to help researchers 
calculate these indices.

## Data: tutorial example

To show a very simple example of the data structure used by selfcomplexity, we
will pull the data of a single participant from the `complexity_data` dataset
provided with the package. This dataset is the same as the one used for the
worked example in our publication - see references below.

```{r}
tutorial_data <- phi_data <- complexity_data[complexity_data$ResponseId == "R_2wmF4JXKNx6McGd",]
dim(tutorial_data)
tutorial_data
```

Note that the first column corresponds to the participant's unique ID, which is
the same in all three rows - we have the card sort from a single participant.
The second column indicates the number of self-aspects that this participant listed.
The third column provides the subtype number. The fourth column corresponds to 
the name of the self-aspect - we've got three different ones here. And the fifth 
column gives us the attributes that were listed
with each self-aspect, separated by commas. 

This long data format is expected by all functions in the package - each row
should correspond to one self-aspect.

## Calculating overlap

The `calculate_overlap` function allows to calculate the overlap index (Rafaeli-Mor et al.,
1999). This index is calculated pairwise - for each pair of aspects, we look
at the number of attributes that they share, and we then sum those across all
pairs and scale for the number of aspects in that participant's card sort. The
mathematical formula is as follows:

$$OL = \frac{\sum_i(\frac{\sum_jC{ij}}{T{j}})}{n(n-1)}$$

where C is the number of attributes common to the two aspects, T is the total 
number of attributes in the referent aspect, n is the total number of aspects 
in the individual’s trait-sort and i and j vary from 0 to n.

The `calculate_overlap` function call requires 5 arguments: the data frame with
the card sort data, the name of the column containing attributes, the name of the
column containing unique IDs, the name of the column containing subtype names,
and whether or not rows with missing subtype names should be removed. The output
is a tibble containing the unique IDs and an overlap index corresponding to each
of those unique IDs. 

```{r}
overlap <- calculate_overlap(data = tutorial_data, att_column = "Attributes",
                             id_column = "ResponseId", subtype_column = "Name",
                             na_name_rm = TRUE)
overlap
```

## Calculating the H statistic

The `calculate_H` function allows to calculate the Linville's (1987) dimensionality
statistic, where a higher score indicates higher complexity. This approach
uses the Scott dimensionality formula (1969):

$$H = \log_{2}(n) - \frac{\sum_{i}n_{i}\log_{2}(n_{i})}{n}$$

where n is the total number of attributes, and $n_i$ is the number of attributes 
that appear in a particular combination in the descriptions of self-aspects. 

The `calculate_H` function call requires 4 arguments: the data frame with
the card sort data, the name of the column containing attributes, the name of the
column containing unique IDs, and a vector containing all attributes that were
made available to the participant during the card sort task. The output
is a tibble containing the unique IDs and an H statistic corresponding to each
of those unique IDs. 

```{r}
data(Attributes_40, package = "selfcomplexity") # loads the vector of attributes
calculate_H(tutorial_data, Attributes, ResponseId, Attributes_40)
```

## Calculating the Phi statistics

Phi is based on a chi‐squared statistic and provides an index of the extent to 
which the distribution of positive and negative traits across self‐aspect groups 
deviates from what would be expected on the basis of chance. 

The `calculate_phi` function call requires 5 arguments: the data frame with
the card sort data, the name of the column containing attributes, the name of the
column containing unique IDs, a vector containing all positive attributes that were
made available to the participant during the card sort task, and a vector 
containing all negative attributes that were made available to the participant 
during the card sort task. The output is a tibble containing the unique IDs and 
a phi statistic corresponding to each of those unique IDs. 

The `selfcomplexity` package includes `Attributes_df`, which is a data frame of 
40 attributes often used with the Showers card sort task (Showers & Zeigler-Hill, 2003),
and an indication of  which of those attributes are positive and negative. 
This data frame can be used
as a basis for extracting the set of positive and negative attributes

```{r}
# preparing the attribute sets
data(Attributes_df, package = "selfcomplexity")

pos <- Attributes_df %>%
dplyr::filter(Positive == 1) %>%
dplyr::pull(Attribute)

neg <- Attributes_df %>%
dplyr::filter(Negative == 1) %>%
dplyr::pull(Attribute)

# function call
calculate_phi(data = tutorial_data, att_column = "Attributes", id_column = "ResponseId", 
              pos_att_vector = pos, neg_att_vector = neg)
```

## Further reading

To read more about the self-complexity indices and work through examples by hand,
see our preprint: [Calculating self-complexity indices, including Linville’s H: Toward standardization and replication.](https://doi.org/10.31234/osf.io/trcwg)

## References
Linville, P. W. (1987). Self-Complexity as a Cognitive Buffer against Stress-Related Illness and Depression. Journal of Personality and Social Psychology, 52(4), 663-676. 

Rafaeli-Mor, E., Gotlib, I. H., & Revelle, W. (1999). The meaning and measurement of self-complexity. Personality and Individual Differences, 27(2), 341-356.

Scott, W. A. (1969). Structure of natural cognitions. Journal of Personality and Social Psychology, 12(4), 261-278.

Showers, C. J., & Zeigler-Hill, V. (2003). Organization of self-knowledge: Features, functions, and flexibility. In Handbook of self and identity. (pp. 47-67). The Guilford Press.
